---
title: "dtw_oecd_2022"
author: "Kefan Yu"
date: "2024-01-05"
output: html_document
---

```{r}
library(readxl)
library(dtwclust)
library(dtw)
```


```{r}
reading <- read_excel("../../data/oecd/oecd_2022_clean.xlsx", sheet = "Reading")
math <- read_excel("../../data/oecd/oecd_2022_clean.xlsx", sheet = "Math")
sci <- read_excel("../../data/oecd/oecd_2022_clean.xlsx", sheet = "Science")
```

```{r}
reading <- subset(reading, select = -c(...1) )
math <- subset(math, select = -c(...1) )
sci <- subset(sci, select = -c(...1) )
```

# Check unattended countries in 2022 (curiousity)

```{r}
reading_missing <- list(year2000 = c(),
                year2003 = c(),
                year2006 = c(),
                year2009 = c(),
                year2012 = c(),
                year2015 = c(),
                year2018 = c(),
                year2022 = c())

reading_nan <- which(is.na(reading$`2000`) | is.na(reading$`2003`) | is.na(reading$`2006`) | is.na(reading$`2009`) | is.na(reading$`2012`) | is.na(reading$`2015`) | is.na(reading$`2018`) | is.na(reading$`2022`))

for (reading_index in reading_nan) {
  
  if (is.na(reading$`2000`[reading_index])){
    reading_missing$year2000 <- append(reading_missing$year2000, reading$Countries[reading_index])
  }
  if (is.na(reading$`2003`[reading_index])){
    reading_missing$year2003 <- append(reading_missing$year2003, reading$Countries[reading_index])
  }
  if (is.na(reading$`2006`[reading_index])){
    reading_missing$year2006 <- append(reading_missing$year2006, reading$Countries[reading_index])
  }
  if (is.na(reading$`2009`[reading_index])){
    reading_missing$year2009 <- append(reading_missing$year2009, reading$Countries[reading_index])
  }
  if (is.na(reading$`2012`[reading_index])){
    reading_missing$year2012 <- append(reading_missing$year2012, reading$Countries[reading_index])
  }
  if (is.na(reading$`2015`[reading_index])){
    reading_missing$year2015 <- append(reading_missing$year2015, reading$Countries[reading_index])
  }
  if (is.na(reading$`2018`[reading_index])){
    reading_missing$year2018 <- append(reading_missing$year2018, reading$Countries[reading_index])
  }
  if (is.na(reading$`2022`[reading_index])){
    reading_missing$year2022 <- append(reading_missing$year2022, reading$Countries[reading_index])
  }
    #cat("Countries: ",reading$Countries[reading_index], "\n")
  
}

print(reading_missing)
```

```{r}
math_missing <- list(year2000 = c(),
                year2003 = c(),
                year2006 = c(),
                year2009 = c(),
                year2012 = c(),
                year2015 = c(),
                year2018 = c(),
                year2022 = c())

math_nan <- which(is.na(math$`2000`) | is.na(math$`2003`) | is.na(math$`2006`) | is.na(math$`2009`) | is.na(math$`2012`) | is.na(math$`2015`) | is.na(math$`2018`))

for (math_index in math_nan) {
  
  if (is.na(math$`2000`[math_index])){
    math_missing$year2000 <- append(math_missing$year2000, math$Countries[math_index])
  }
  if (is.na(math$`2003`[math_index])){
    math_missing$year2003 <- append(math_missing$year2003, math$Countries[math_index])
  }
  if (is.na(math$`2006`[math_index])){
    math_missing$year2006 <- append(math_missing$year2006, math$Countries[math_index])
  }
  if (is.na(math$`2009`[math_index])){
    math_missing$year2009 <- append(math_missing$year2009, math$Countries[math_index])
  }
  if (is.na(math$`2012`[math_index])){
    math_missing$year2012 <- append(math_missing$year2012, math$Countries[math_index])
  }
  if (is.na(math$`2015`[math_index])){
    math_missing$year2015 <- append(math_missing$year2015, math$Countries[math_index])
  }
  if (is.na(math$`2018`[math_index])){
    math_missing$year2018 <- append(math_missing$year2018, math$Countries[math_index])
  }
  
  if (is.na(math$`2022`[math_index])){
    math_missing$year2022 <- append(math_missing$year2022, math$Countries[math_index])
  }
    #cat("Countries: ",reading$Countries[reading_index], "\n")
  
}

print(math_missing)
```

```{r}
sci_missing <- list(year2000 = c(),
                year2003 = c(),
                year2006 = c(),
                year2009 = c(),
                year2012 = c(),
                year2015 = c(),
                year2018 = c(),
                year2022 = c())

sci_nan <- which(is.na(sci$`2000`) | is.na(sci$`2003`) | is.na(sci$`2006`) | is.na(sci$`2009`) | is.na(sci$`2012`) | is.na(sci$`2015`) | is.na(sci$`2018`))

for (sci_index in sci_nan) {
  
  if (is.na(sci$`2000`[sci_index])){
    sci_missing$year2000 <- append(sci_missing$year2000, sci$Countries[sci_index])
  }
  if (is.na(sci$`2003`[sci_index])){
    sci_missing$year2003 <- append(sci_missing$year2003, sci$Countries[sci_index])
  }
  if (is.na(sci$`2006`[sci_index])){
    sci_missing$year2006 <- append(sci_missing$year2006, sci$Countries[sci_index])
  }
  if (is.na(sci$`2009`[sci_index])){
    sci_missing$year2009 <- append(sci_missing$year2009, sci$Countries[sci_index])
  }
  if (is.na(sci$`2012`[sci_index])){
    sci_missing$year2012 <- append(sci_missing$year2012, sci$Countries[sci_index])
  }
  if (is.na(sci$`2015`[sci_index])){
    sci_missing$year2015 <- append(sci_missing$year2015, sci$Countries[sci_index])
  }
  if (is.na(sci$`2018`[sci_index])){
    sci_missing$year2018 <- append(sci_missing$year2018, sci$Countries[sci_index])
  }
  if (is.na(sci$`2022`[sci_index])){
    sci_missing$year2022 <- append(sci_missing$year2022, sci$Countries[sci_index])
  }
    #cat("Countries: ",reading$Countries[reading_index], "\n")
  
}
print(sci_missing)
```

From above, we can see that the only country that miss the 2022 exam is Luxembourg.

# Data Cleaning

```{r}
country = reading[['Countries']]

dtw_reading <- reading[, -1]

dtw_reading
```

```{r}
library(zoo)
dtw_reading[] <- t(apply(dtw_reading, 1, function(x) na.locf0(na.locf0(x), fromLast = TRUE)))
dtw_reading
```
# Reading 

```{r}
reading_lst <- tslist(dtw_reading)

reading_cvi <- list()
for (i in 2:10){
  reading_clust <- tsclust(reading_lst, type = "partitional", k = i, distance = "dtw", centroid = "pam",set.seed(5))
  reading_metric <- cvi(reading_clust, type = "valid", log.base = 10)
  reading_cvi <- append(reading_cvi, list(reading_metric))
}

reading_cvi_ma <- do.call(rbind, reading_cvi)
rw <- c("K2","K3","K4","K5","K6","K7","K8","K9","K10")
rownames(reading_cvi_ma) <- rw
print(reading_cvi_ma)
```

```{r}
reading_clust_opt <- tsclust(reading_lst, type = "partitional", k = 4, distance = "dtw", centroid = "pam",seed = 725)
plot(reading_clust_opt)
```

```{r}
print(t(cbind(reading$Countries, cluster = reading_clust_opt@cluster)))
```

```{r}
result_reading_avg = cbind(reading$Countries, cluster = reading_clust_opt@cluster)
result_reading_avg <- as.data.frame(result_reading_avg)

cat("Number of countries in K1: ", nrow(result_reading_avg[result_reading_avg['cluster'] == "1", ]), "\n")
cat("Number of countries in K2: ", nrow(result_reading_avg[result_reading_avg['cluster'] == "2", ]), "\n")
cat("Number of countries in K3: ", nrow(result_reading_avg[result_reading_avg['cluster'] == "3", ]), "\n")
cat("Number of countries in K4: ", nrow(result_reading_avg[result_reading_avg['cluster'] == "4", ]), "\n")
```

```{r}
reading_clust_opt <- tsclust(reading_lst, type = "partitional", k = 4, distance = "dtw", centroid = "pam",seed = 725)

result_reading_avg = cbind(reading$Countries, cluster = reading_clust_opt@cluster)
result_reading_avg <- as.data.frame(result_reading_avg)

cat("Countries in K1: ", result_reading_avg$V1[result_reading_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_reading_avg$V1[result_reading_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_reading_avg$V1[result_reading_avg$cluster == "3"], "\n")
cat("Countries in K4: ", result_reading_avg$V1[result_reading_avg$cluster == "4"], "\n")
```

```{r}
reading_clust_opt2 <- tsclust(reading_lst, type = "partitional", k = 4, distance = "dtw", centroid = "pam",seed = 825)

result_reading_avg2 = cbind(reading$Countries, cluster = reading_clust_opt2@cluster)
result_reading_avg2 <- as.data.frame(result_reading_avg2)

cat("Countries in K1: ", result_reading_avg2$V1[result_reading_avg2$cluster == "1"], "\n")
cat("Countries in K2: ", result_reading_avg2$V1[result_reading_avg2$cluster == "2"], "\n")
cat("Countries in K3: ", result_reading_avg2$V1[result_reading_avg2$cluster == "3"], "\n")
cat("Countries in K4: ", result_reading_avg2$V1[result_reading_avg2$cluster == "4"], "\n")
```

```{r}
reading_clust_opt3 <- tsclust(reading_lst, type = "partitional", k = 4, distance = "dtw", centroid = "pam",seed = 925)

result_reading_avg3 = cbind(reading$Countries, cluster = reading_clust_opt3@cluster)
result_reading_avg3 <- as.data.frame(result_reading_avg3)

cat("Countries in K1: ", result_reading_avg3$V1[result_reading_avg3$cluster == "1"], "\n")
cat("Countries in K2: ", result_reading_avg3$V1[result_reading_avg3$cluster == "2"], "\n")
cat("Countries in K3: ", result_reading_avg3$V1[result_reading_avg3$cluster == "3"], "\n")
cat("Countries in K4: ", result_reading_avg3$V1[result_reading_avg3$cluster == "4"], "\n")
```

```{r}
reading_clust_opt4 <- tsclust(reading_lst, type = "partitional", k = 4, distance = "dtw", centroid = "pam",seed = 1025)

result_reading_avg4 = cbind(reading$Countries, cluster = reading_clust_opt4@cluster)
result_reading_avg4 <- as.data.frame(result_reading_avg4)

cat("Countries in K1: ", result_reading_avg4$V1[result_reading_avg4$cluster == "1"], "\n")
cat("Countries in K2: ", result_reading_avg4$V1[result_reading_avg4$cluster == "2"], "\n")
cat("Countries in K3: ", result_reading_avg4$V1[result_reading_avg4$cluster == "3"], "\n")
cat("Countries in K4: ", result_reading_avg4$V1[result_reading_avg4$cluster == "4"], "\n")
```

```{r}
reading_clust_opt5 <- tsclust(reading_lst, type = "partitional", k = 4, distance = "dtw", centroid = "pam",seed = 1125)

result_reading_avg5 = cbind(reading$Countries, cluster = reading_clust_opt5@cluster)
result_reading_avg5 <- as.data.frame(result_reading_avg5)

cat("Countries in K1: ", result_reading_avg5$V1[result_reading_avg5$cluster == "1"], "\n")
cat("Countries in K2: ", result_reading_avg5$V1[result_reading_avg5$cluster == "2"], "\n")
cat("Countries in K3: ", result_reading_avg5$V1[result_reading_avg5$cluster == "3"], "\n")
cat("Countries in K4: ", result_reading_avg5$V1[result_reading_avg5$cluster == "4"], "\n")
```

By trying different seeds, we can see that the result is very unstable. However, countries like Chile, Colombia, Costa Rica, Mexico, Türkiye always tend to cluster together, and seed 1025 and 1125 both have its group. Thus we choose seed 1025 for dtwclust. Withy the 2022 scores, Chile Colombia and Mexico also tend to cluster together.

```{r}
plot(reading_clust_opt4)
```

```{r}
reading_clust_opt_h <- tsclust(reading_lst, type = "h", k = 4, distance = "dtw", seed = 725)
plot(reading_clust_opt_h, type = "sc")
```

```{r}
#print(t(cbind(reading$Countries, cluster = reading_clust_opt_h@cluster)))

result_reading_avg_h <- as.data.frame(cbind(reading$Countries, cluster = reading_clust_opt_h@cluster))

cat("Countries in K1: ", result_reading_avg_h$V1[result_reading_avg_h$cluster == "1"], "\n")
cat("Countries in K2: ", result_reading_avg_h$V1[result_reading_avg_h$cluster == "2"], "\n")
cat("Countries in K3: ", result_reading_avg_h$V1[result_reading_avg_h$cluster == "3"], "\n")
cat("Countries in K4: ", result_reading_avg_h$V1[result_reading_avg_h$cluster == "4"], "\n")
```

```{r}
result_reading_avg_h = cbind(reading$Countries, cluster = reading_clust_opt_h@cluster)
result_reading_avg_h <- as.data.frame(result_reading_avg_h)

cat("Number of countries in K1: ", nrow(result_reading_avg_h[result_reading_avg_h['cluster'] == "1", ]), "\n")
cat("Number of countries in K2: ", nrow(result_reading_avg_h[result_reading_avg_h['cluster'] == "2", ]), "\n")
cat("Number of countries in K3: ", nrow(result_reading_avg_h[result_reading_avg_h['cluster'] == "3", ]), "\n")
cat("Number of countries in K4: ", nrow(result_reading_avg_h[result_reading_avg_h['cluster'] == "4", ]), "\n")
```

By using the hierarchical clustering, we can see that Chile and Colombia are not in the same group but Colomibia and Mexico are still in the same group. I wonder it might be the reason that Colombia and Mexico have very similar scores since they are the least and the second least ranking among all oECD countries. Thus only they are in the same group. It is also notable to see that K2 has 25 countries that share similar trends.

# Math

```{r}
dtw_math <- math[, -1]

dtw_math <- subset(dtw_math, select = -c(`2000`) )
```

```{r}
dtw_math[] <- t(apply(dtw_math, 1, function(x) na.locf0(na.locf0(x), fromLast = TRUE)))
```

```{r}
math_lst <- tslist(dtw_math)
```

```{r}
math_cvi <- list()
for (i in 2:10){
  math_clust <- tsclust(math_lst, type = "partitional", k = i, distance = "dtw", centroid = "pam",set.seed(4))
  math_metric <- cvi(math_clust, type = "valid", log.base = 10)
  math_cvi <- append(math_cvi, list(math_metric))
}

math_cvi_ma <- do.call(rbind, math_cvi)
rownames(math_cvi_ma) <- rw
print(math_cvi_ma)
```
We can see that by adding 2022 scores, K=4 is not a good choice anymore, the Silhoutte value ia smaller than K=3's Sil value, and DB is larger, and there is no significant difference between D values. Thus I choose K = 3.

```{r}
math_clust_opt <- tsclust(math_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 725)

result_math_avg = cbind(reading$Countries, cluster = math_clust_opt@cluster)
result_math_avg <- as.data.frame(result_math_avg)

cat("Countries in K1: ", result_math_avg$V1[result_math_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_math_avg$V1[result_math_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_math_avg$V1[result_math_avg$cluster == "3"], "\n")
```

```{r}
math_clust_opt2 <- tsclust(math_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 825)

result_math_avg2 = cbind(reading$Countries, cluster = math_clust_opt2@cluster)
result_math_avg2 <- as.data.frame(result_math_avg2)

cat("Countries in K1: ", result_math_avg2$V1[result_math_avg2$cluster == "1"], "\n")
cat("Countries in K2: ", result_math_avg2$V1[result_math_avg2$cluster == "2"], "\n")
cat("Countries in K3: ", result_math_avg2$V1[result_math_avg2$cluster == "3"], "\n")
```

```{r}
math_clust_opt3 <- tsclust(math_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 925)

result_math_avg3 = cbind(reading$Countries, cluster = math_clust_opt3@cluster)
result_math_avg3 <- as.data.frame(result_math_avg3)

cat("Countries in K1: ", result_math_avg3$V1[result_math_avg3$cluster == "1"], "\n")
cat("Countries in K2: ", result_math_avg3$V1[result_math_avg3$cluster == "2"], "\n")
cat("Countries in K3: ", result_math_avg3$V1[result_math_avg3$cluster == "3"], "\n")
```

Since K only equals to 3, there is no variation among different seeds. We can see that Chile, Colombia, Costa Rica, Mexico, Türkiye form a single group.

```{r}

plot(math_clust_opt)

```

When K is 3, the clustering results are usually based on the score range (like the sci result). From the graph above, we can see that it is divided into 450-500 (middle score), 380-440(low score), 475 - 550 (high score).

```{r}
math_clust_opt_h <- tsclust(math_lst, type = "h", k = 3, distance = "dtw", seed = 725)
plot(math_clust_opt_h, type="sc")
```

```{r}
result_math_avg_h = cbind(reading$Countries, cluster = math_clust_opt_h@cluster)
result_math_avg_h <- as.data.frame(result_math_avg_h)

cat("Countries in K1: ", result_math_avg_h$V1[result_math_avg_h$cluster == "1"], "\n")
cat("Countries in K2: ", result_math_avg_h$V1[result_math_avg_h$cluster == "2"], "\n")
cat("Countries in K3: ", result_math_avg_h$V1[result_math_avg_h$cluster == "3"], "\n")
```

The hierarchical clustering throw every country except Chile, Colombia, Costa Rica, Mexico, Türkiye, Greece and Israel into one cluster, and divide them into two clusters. Such clustering is not based on the similarity of each sequence, but based on their score ranges.

# Science

```{r}
dtw_sci <- sci[, -1]
dtw_sci <- subset(dtw_sci, select = -c(`2000`,`2003`) )
dtw_sci[] <- t(apply(dtw_sci, 1, function(x) na.locf0(na.locf0(x), fromLast = TRUE)))
sci_lst <- tslist(dtw_sci)


sci_cvi <- list()
for (i in 2:10){
  sci_clust <- tsclust(sci_lst, type = "partitional", k = i, distance = "dtw", centroid = "pam",set.seed(4))
  sci_metric <- cvi(sci_clust, type = "valid", log.base = 10)
  sci_cvi <- append(sci_cvi, list(sci_metric))
}

sci_cvi_ma <- do.call(rbind, sci_cvi)
rownames(sci_cvi_ma) <- rw
print(sci_cvi_ma)

```

Similar as the math situation, K=3 is better in this case

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 725)
result_sci_avg = cbind(reading$Countries, cluster = sci_clust_opt@cluster)
result_sci_avg <- as.data.frame(result_sci_avg)

cat("Countries in K1: ", result_sci_avg$V1[result_sci_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_sci_avg$V1[result_sci_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_sci_avg$V1[result_sci_avg$cluster == "3"], "\n")
```

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 825)
result_sci_avg = cbind(reading$Countries, cluster = sci_clust_opt@cluster)
result_sci_avg <- as.data.frame(result_sci_avg)

cat("Countries in K1: ", result_sci_avg$V1[result_sci_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_sci_avg$V1[result_sci_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_sci_avg$V1[result_sci_avg$cluster == "3"], "\n")
```

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 925)
result_sci_avg = cbind(reading$Countries, cluster = sci_clust_opt@cluster)
result_sci_avg <- as.data.frame(result_sci_avg)

cat("Countries in K1: ", result_sci_avg$V1[result_sci_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_sci_avg$V1[result_sci_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_sci_avg$V1[result_sci_avg$cluster == "3"], "\n")
```

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 1025)
result_sci_avg = cbind(reading$Countries, cluster = sci_clust_opt@cluster)
result_sci_avg <- as.data.frame(result_sci_avg)

cat("Countries in K1: ", result_sci_avg$V1[result_sci_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_sci_avg$V1[result_sci_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_sci_avg$V1[result_sci_avg$cluster == "3"], "\n")
```

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 1125)
result_sci_avg = cbind(reading$Countries, cluster = sci_clust_opt@cluster)
result_sci_avg <- as.data.frame(result_sci_avg)

cat("Countries in K1: ", result_sci_avg$V1[result_sci_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_sci_avg$V1[result_sci_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_sci_avg$V1[result_sci_avg$cluster == "3"], "\n")
```

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 3, distance = "dtw", centroid = "pam",seed = 1225)
result_sci_avg = cbind(reading$Countries, cluster = sci_clust_opt@cluster)
result_sci_avg <- as.data.frame(result_sci_avg)

cat("Countries in K1: ", result_sci_avg$V1[result_sci_avg$cluster == "1"], "\n")
cat("Countries in K2: ", result_sci_avg$V1[result_sci_avg$cluster == "2"], "\n")
cat("Countries in K3: ", result_sci_avg$V1[result_sci_avg$cluster == "3"], "\n")
```

Since seed 825, 925, 1225 converge, we use 1225 for analysis.

```{r}
plot(sci_clust_opt)
```

Similar for math, the groups are generated based on score range. Thus it makes sense that there are many countries in the middle score cluster.

```{r}
sci_clust_opt_h <- tsclust(sci_lst, type = "h", k = 3, distance = "dtw", seed = 725)
plot(sci_clust_opt_h, type="sc")
```

```{r}
print(t(cbind(reading$Countries, cluster = sci_clust_opt_h@cluster)))
```