---
title: "dtw_cluster"
author: "Kefan Yu"
date: "2023-10-10"
output: html_document
---

```{r}
library(dtwclust)
library(dtw)
```

```{r}
reading <- read.csv("../data_all/dtw_reading.csv")
head(reading)
```


```{r}
reading <- subset(reading, select = -c(X) )
reading
```

```{r}
colnames(reading)[colnames(reading) == "X2000"] = "2000"
colnames(reading)[colnames(reading) == "X2003"] = "2003"
colnames(reading)[colnames(reading) == "X2006"] = "2006"
colnames(reading)[colnames(reading) == "X2009"] = "2009"
colnames(reading)[colnames(reading) == "X2012"] = "2012"
colnames(reading)[colnames(reading) == "X2015"] = "2015"
colnames(reading)[colnames(reading) == "X2018"] = "2018"

reading
```
```{r}
country = reading[['Countries']]
```

```{r}
dtw_reading <- reading[, -1]
```

```{r}
reading_lst <- tslist(dtw_reading)
```

```{r}
remove_nan <- function(ts) {
  ts[!is.na(ts)]
}

# Apply the function to each time series in the list
reading_lst <- lapply(reading_lst, remove_nan)
```

```{r}
reading_cvi <- list()
for (i in 2:10){
  reading_clust <- tsclust(reading_lst, type = "partitional", k = i, distance = "dtw_basic", centroid = "pam")
  reading_metric <- cvi(reading_clust, type = "valid", log.base = 10)
  reading_cvi <- append(reading_cvi, list(reading_metric))
}
```

```{r}
reading_cvi_ma <- do.call(rbind, reading_cvi)
rw <- c("K2","K3","K4","K5","K6","K7","K8","K9","K10")
rownames(reading_cvi_ma) <- rw
print(reading_cvi_ma)
```
– "Sil" (!): Silhouette index (Rousseeuw (1987); to be maximized).
– "SF" (~): Score Function (Saitta et al. (2007); to be maximized; see notes).
– "CH" (~): Calinski-Harabasz index (Arbelaitz et al. (2013); to be maximized).
– "DB" (?): Davies-Bouldin index (Arbelaitz et al. (2013); to be minimized).
– "DBstar" (?): Modified Davies-Bouldin index (DB*) (Kim and Ramakrishna (2005); to
be minimized).
– "D" (!): Dunn index (Arbelaitz et al. (2013); to be maximized).
– "COP" (!): COP index (Arbelaitz et al. (2013); to be minimized).

The optimal cluster numbers would be 6

```{r}
reading_clust_opt <- tsclust(reading_lst, type = "partitional", k = 6, distance = "dtw", centroid = "pam",seed = 725)
plot(reading_clust_opt)
```

```{r}
print(t(cbind(reading$Countries, cluster = reading_clust_opt@cluster)))
```
MATH DTW

```{r}
math <- read.csv("../data_all/dtw_math.csv")
math <- subset(math, select = -c(X) )

colnames(math)[colnames(math) == "X2000"] = "2000"
colnames(math)[colnames(math) == "X2003"] = "2003"
colnames(math)[colnames(math) == "X2006"] = "2006"
colnames(math)[colnames(math) == "X2009"] = "2009"
colnames(math)[colnames(math) == "X2012"] = "2012"
colnames(math)[colnames(math) == "X2015"] = "2015"
colnames(math)[colnames(math) == "X2018"] = "2018"

dtw_math <- math[, -1]

```

```{r}
math_lst <- tslist(dtw_math)

# Apply the function to each time series in the list
math_lst <- lapply(math_lst, remove_nan)

print(math_lst)
```
```{r}
math_cvi <- list()
for (i in 2:10){
  math_clust <- tsclust(math_lst, type = "partitional", k = i, distance = "dtw", centroid = "pam")
  math_metric <- cvi(math_clust, type = "valid", log.base = 10)
  math_cvi <- append(math_cvi, list(math_metric))
}

print(math_cvi)
```

```{r}
math_cvi_ma <- do.call(rbind, math_cvi)
rownames(math_cvi_ma) <- rw
print(math_cvi_ma)
```

```{r}
math_clust_opt <- tsclust(math_lst, type = "partitional", k = 7, distance = "dtw", centroid = "pam",seed = 725)
plot(math_clust_opt)
```

```{r}
print(t(cbind(math$Countries, cluster = math_clust_opt@cluster)))
```

```{r}
sci <- read.csv("../data_all/dtw_sci.csv")
sci <- subset(sci, select = -c(X) )

colnames(sci)[colnames(sci) == "X2000"] = "2000"
colnames(sci)[colnames(sci) == "X2003"] = "2003"
colnames(sci)[colnames(sci) == "X2006"] = "2006"
colnames(sci)[colnames(sci) == "X2009"] = "2009"
colnames(sci)[colnames(sci) == "X2012"] = "2012"
colnames(sci)[colnames(sci) == "X2015"] = "2015"
colnames(sci)[colnames(sci) == "X2018"] = "2018"

dtw_sci <- sci[, -1]
```

```{r}
sci_lst <- tslist(dtw_sci)

# Apply the function to each time series in the list
sci_lst <- lapply(math_lst, remove_nan)

```

```{r}
sci_cvi <- list()
for (i in 2:10){
  sci_clust <- tsclust(sci_lst, type = "partitional", k = i, distance = "dtw", centroid = "pam")
  sci_metric <- cvi(sci_clust, type = "valid", log.base = 10)
  sci_cvi <- append(sci_cvi, list(sci_metric))
}

sci_cvi_ma <- do.call(rbind, sci_cvi)
rownames(sci_cvi_ma) <- rw
print(sci_cvi_ma)
```

```{r}
sci_clust_opt <- tsclust(sci_lst, type = "partitional", k = 6, distance = "dtw", centroid = "pam",seed = 725)
plot(sci_clust_opt)
```

```{r}
print(t(cbind(sci$Countries, cluster = sci_clust_opt@cluster)))
```